Supporting Information

### 2018-03-30 SupportingInformation.Rmd
finished editing up through aphid count

Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a permutation 
of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as those of the original data. 
Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
will be close to $\tau$ (differences arising only through sampling variation). If the instead of $\theta_\tau$
a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
sampling variation) the Spearman correlation of the empirical data.

The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the empirical data, to the distribution
of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
provides a test of whether dependence in the part of the distribution specified by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but unchanged marginals and unchanged 
Kendall (or Spearman)
correlation. Significant deviations correspond to deviations from normal copula structure. 
The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
provides a test of asymmetry of tail dependence.

For multivariate datasets, e.g., the aphid abundance or first flight data, each of which consists of time series 
from 10 locations, a spatial resampling scheme was used instead of the surrogate approach described above.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 as part of a method for 
performing model selection among matrix regression models. For each pair of locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The values $s_{i,j}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a distribution of bootstrapped values.
Bootstrapping was performed by resampling the locations of measurement, with replacement; this corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
@bjornstadfalck2001, @walteretal2017.

@@

As described in Methods, each of our statistics, as computed for the 
ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
of values of the same statistic with the same ranges computed on 1000 normal-copula 
surrogate datasets with similar Kendall or Spearman correlations, in different 
runs for Kendall and Spearman (Figure \ref{fig_soilCN_nonparam}). Because the correlation 
used in the $\cor$ statistic is the Spearman correlation, Spearman-preserving surrogates
should be used, so the top-right panel of Figure \ref{fig_soilCN_nonparam}
is more reliable than the top left. Results indicate that tail dependence of 
data is stronger in both the lower and upper tails than would be expected under a 
null hypothesis of a normal copula. The dependence between C and N is also 
weaker in the middle of the distributions
than would be expected based on a normal-copula null model.

@@

We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_soilCNcop}. 
Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data.

@@

```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("NonParamStat.R"))}
set.seed(seed) 
resloc<-"./Results/stat_results/stat_aphid_count/"

@@

```{r tab_count_resamp,echo=F,results="markup"}
s<-stat_aphid_count$numericdf[,2:5]
rownames(s)<-c("spear","kend","cor~l~","cor~u~","P~l~","P~u~","$D^2$~l~","$D^2$~u~","cor~l~ - cor~u~","P~l~ - P~u~","$D^2$~u~ - $D^2$~l~")
knitr::kable(s,
             format = "pandoc",
             caption = "Aphid_count : resampling result for stat indicating lower tail dependence\\label{tab_count_resamp}",
             booktabs=TRUE)
```

![copula plot for aphid-count\label{fig_count_allcop}](./Results/stat_results/stat_aphid_count/Sp_10_AllCops.pdf)

![Plot for aphid-count : Spearman vs. Distance\label{fig_count_spear}](./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf){ width=50%, height=50% }

@@

Table \ref{tab_count_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean. Fourth column named $fracgt0$ in Table \ref{tab_count_resamp} indicates the fraction of resampled mean greater then zero. $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ statistic with positive mean and almost $100$% $fracgt0$ proves that *Green spruce aphid*-count data has significant lower tail dependence.

>>>>

Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a permutation 
of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as those of the original data. 
Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
sampling variation) the Spearman correlation of the empirical data.

The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the empirical data, to the distribution
of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
provides a test of whether dependence in the part of the distribution specified by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but unchanged marginals and unchanged 
Kendall (or Spearman)
correlation. Significant deviations correspond to deviations from normal copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different from that of a normal copula.  
The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
provides a test of asymmetry of tail dependence.

For multivariate datasets, e.g., the aphid abundance or first flight data, each of which consists of time series 
from 10 locations, a spatial resampling scheme was used instead of the surrogate approach described above.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 as part of a method for 
performing model selection among matrix regression models. For each pair of locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The values $s_{i,j}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a distribution of bootstrapped values.
Bootstrapping was performed by resampling the locations of measurement, with replacement; this corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
@bjornstadfalck2001, @walteretal2017. 
<!--Shya, these citations render as (4), (5), whereas they should render as (4, 5), or better still, (Bjornstad & Falck, 2001; Walter et al, 2017) (the author-date callout style is better when space is not a limitation because it gives the reader more information without having to look in the bibliography). Can you fix this, please?-->

@@

As described in Methods, each of our statistics, as computed for the 
ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
of values of the same statistic with the same ranges computed on 1000 normal-copula 
surrogate datasets with similar Kendall or Spearman correlations, in different 
runs for Kendall and Spearman (Figure \ref{fig_soilCN_nonparam}). Because the correlation 
used in the $\cor$ statistic is the Spearman correlation, Spearman-preserving surrogates
should be used, so the top-right panel of Figure \ref{fig_soilCN_nonparam}
is more reliable than the top left. Results indicate that tail dependence of 
data is stronger in both the lower and upper tails than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
weaker in the middle of the distributions
than would be expected based on a normal-copula null model.

@@

We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_soilCNcop}. 
Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.

@@

<!--Do the main computations for this section-->
```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("NonParamStat.R"))}
set.seed(seed) 
resloc<-"./Results/stat_results/stat_aphid_count/"

@@

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
resampling scheme, as described in Methods (Table \ref{tab_count_resamp}). Confidence intervals of
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
structure do not have, providing further evidence answering the third part of question 1.

<!--Show the statistical results in a table-->
```{r tab_count_resamp,echo=F,results="markup"}
s<-stat_aphid_count$numericdf[,c(3,2,4)]
colnames(s)<-c("Statistic","Lower 95\\% CI", "Mean", "Upper 95\\% CI")
rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Psq_l - \\Psq_u$","$\\Dsq_u - \\Dsq_l$")
knitr::kable(s,
             format = "pandoc",
             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for abundances of the green spruce aphid.\\label{tab_count_resamp}",
             booktabs=TRUE)
```

<!--Shya, I dont think we need this, so I have commented it out. If you disagree, let me know, otherwise
please go ahead and delete it and this comment.
![copula plot for aphid-count\label{fig_count_allcop}](./Results/stat_results/stat_aphid_count/Sp_10_AllCops.pdf)
-->

<!--Shya, can you please organize these following plots into one plot with mutliple panels?-->
![Plot for aphid-count : Spearman vs. Distance\label{fig_count_spear}](./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf){ width=50%, height=50% }

@@
